---
output:
  md_document:
    variant: markdown_github
---

# Purpose

This is the final README, it is a compilation of questions one through seven. 


```{r}

rm(list = ls()) # Clean your environment:
gc() # garbage collection - It can be useful to call gc after a large object has been removed, as this may prompt R to return memory to the operating system.
library(tidyverse)
list.files('code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))
```

```{r setup, eval=FALSE}
#Loading all relevant packages. 
#install.packages("pacman")
pacman::p_load(tidyverse)
pacman::p_load(readxl)
pacman::p_load(readr)
pacman::p_load(ggplot2)
pacman::p_load(gt)
pacman::p_load(dplyr)
library(gt)
library(tidyverse)
library(dplyr)
```


```{r}
#Loading all relevant code from each questions code folder. 
list.files('C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/Question 1/code', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))
list.files('C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/Question 2/code', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))
list.files('C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/Question 3/code', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))
list.files('C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/Question 4/code', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))
list.files('C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/Question 5/code', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))
list.files('C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/Question 6/code', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))
list.files('C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/Question 7/code', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))
```

# Question 1: Systematic AI fund

```{r}
#Loading relevant data for question 1
library(tidyverse)
ASISA <- readRDS("C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/data/ASISA_Rets.rds") #The ASISA fund data starts on 2003-10-31 and ends 2023-09-30

BM <- readRDS("C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/data/Capped_SWIX.rds")#The BM data starts on 2002-02-28 and ends 2023-09-30

AI_Fund <- readRDS("C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/data/AI_Max_Fund.rds")#The AI_Fund data starts on 2003-01-31 and ends 2023-09-30 

#All the data ends on the same date but starts on different dates. I need to keep this in mind find plotting comparative plots. 
```

## Rolling returns camparison

```{r}
#First loading the relevant packages 
pacman::p_load("xts", "tidyverse", "tbl2xts", "PerformanceAnalytics", 
    "lubridate", "glue", "RcppRoll")

#Now I need to merge the ASISA, BM and AI_fund datasets to ensure comparative analysis is easier. 
library(tidyverse)
library(lubridate)

# Renaming columns for ease
names(AI_Fund)[2] <- "AI_Fund_Return"
names(BM)[3] <- "BM_Return"

# Merging AI Fund and Benchmark Data
combined_returns_data <- left_join(AI_Fund, BM, by = "date")

# Aggregate ASISA Data
ASISA_aggregated <- ASISA %>%
  group_by(date) %>%
  summarize(ASISA_Return = mean(Returns))

# Merge with ASISA Data
final_merged_returns_data <- left_join(combined_returns_data, ASISA_aggregated, by = "date")

# Checking the final merged data
head(final_merged_returns_data)

#Now I can start calculating the rolling returns

library(tidyverse)
library(RcppRoll)
library(fmxdat)


# Prepare the data
Rolling_returns_plot_df <- final_merged_returns_data %>%
  mutate(Rolling_Return_AI_Fund = RcppRoll::roll_mean(AI_Fund_Return, 12, fill = NA, align = "right"),
         Rolling_Return_BM = RcppRoll::roll_mean(BM_Return, 12, fill = NA, align = "right"),
         Rolling_Return_ASISA = RcppRoll::roll_mean(ASISA_Return, 12, fill = NA, align = "right")) %>%
  group_by(date) %>% 
  filter(any(!is.na(Rolling_Return_AI_Fund) | !is.na(Rolling_Return_BM) | !is.na(Rolling_Return_ASISA)))

# Create the plot
Rolling_returns_plot <- Rolling_returns_plot_df %>%
  ggplot() +
  geom_line(aes(date, Rolling_Return_AI_Fund, color = "AI Fund"), alpha = 0.7, size = 1.25) +
  geom_line(aes(date, Rolling_Return_BM, color = "Benchmark"), alpha = 0.7, size = 1.25) +
  geom_line(aes(date, Rolling_Return_ASISA, color = "ASISA"), alpha = 0.7, size = 1.25) +
  labs(title = "Rolling Returns: AI Fund vs Benchmark vs ASISA", x = "", y = "Rolling Return", caption = "Note: Rolling returns calculated over 12 months.") +
  theme_fmx(title.size = ggpts(30), subtitle.size = ggpts(5), caption.size = ggpts(25), CustomCaption = T) +
  fmx_cols()

# Apply finplot
finplot(Rolling_returns_plot, x.date.dist = "1 year", x.date.type = "%Y", x.vert = T, y.pct = T, y.pct_acc = 1)


```
## Making use of the PerformanceAnalytics package to do inference on the ASISA, BM and AI_Fund

```{r}
library(tbl2xts)
library(tidyverse)
library(tidyr)
library(dplyr)
library(PerformanceAnalytics)

#Dropping the Tickets column from the dataset

final_merged_returns_data <- select(final_merged_returns_data, -Tickers)

#Transforming to xts

xts_final_merged_returns_data <-final_merged_returns_data %>% 
    gather(funds, returns, -date) %>% 
    group_by(funds) %>% 
    tbl_xts(tblData = ., cols_to_xts = returns, spread_by = funds)

```

### Some intresting stats 

```{r}
#Now that the merged returns data is in xts I can do some nice analysis 

#Calculate annualized returns for each dataset
ai_fund_stats <- table.AnnualizedReturns(xts_final_merged_returns_data$AI_Fund_Return, scale = 252)
bm_stats <- table.AnnualizedReturns(xts_final_merged_returns_data$BM_Return, scale = 252)
asisa_stats <- table.AnnualizedReturns(xts_final_merged_returns_data$ASISA_Return, scale = 252)
```

My fund = My fund has a very high annual return of 21% which is higher than both the benchmark capped SWIX and the actively managed ASISA. My fund also has an extremely low standard deviation of 0.726% which means that my fund holds very little risk. Furthermore, the Sharpe ratio of 28.9 is extraordinarily high. 

The benchmark fund = The benchmark fund has very decent annual returns of 14.43%. Furthermore the benchmark as expected has a low annualized standard deviation of 0.69%. Thus, meaning the benchmark holds very little risk. Once again, the Sharpe ration of 19.52 is extraordinarily high. 

The active managers = The actice managers have on average an annualized return of 8.1%, this is fairly good but is lagging behind my fund at 21% and the benchmark at 14.43%. 

### Some more intresting stats 

```{r}
AI_Fund_all_stats <- table.DownsideRisk(xts_final_merged_returns_data$AI_Fund_Return)
BM_all_stats <- table.DownsideRisk(xts_final_merged_returns_data$BM_Return)
ASISA_all_stats <- table.DownsideRisk(xts_final_merged_returns_data$ASISA_Return)
```


### Some histogram plots

```{r}

# Add normal line and VaR estimates
AI_Fund_histogram <- chart.Histogram(xts_final_merged_returns_data$AI_Fund_Return, 
  methods = c("add.density", "add.normal", "add.risk"),
  main = "Adding Value at Risk (95%)")


BM_histogram <- chart.Histogram(xts_final_merged_returns_data$BM_Return, 
  methods = c("add.density", "add.normal", "add.risk"),
  main = "Adding Value at Risk (95%)")


ASISA_histogram <- chart.Histogram(xts_final_merged_returns_data$ASISA_Return, 
  methods = c("add.density", "add.normal", "add.risk"),
  main = "Adding Value at Risk (95%)")

```

### Lets now look at a plot that looks at the drawdowns present in each fund, which will evaluate the risk involved in each. 

```{r}
AI_Fund_Drawdowns <- chart.Drawdown(xts_final_merged_returns_data$AI_Fund_Return,main = "Drawdowns: AI_Fund", 
               col = "steelblue")

BM_Drawdowns <- chart.Drawdown(xts_final_merged_returns_data$BM_Return,main = "Drawdowns: Benchmark", 
               col = "steelblue")

ASISA_Drawdowns <- chart.Drawdown(xts_final_merged_returns_data$ASISA_Return,main = "Drawdowns: ASISA", 
               col = "steelblue")
```

AS can be seen from the above drawdowns plots, my AI_Fund and the benchmark follows a very similar risk to drawdowns. On the contrary ASISA (Which pertains to the active managers) contains heavier drawdowns. Thus the active managers contains higher risk than that of the benchmark fund (The market) and my fund 



# Question 2







# Question 3: Portfolio construction

## Loading relevant data and packages 

```{r}
#Relevant data
ALSI <- readRDS("C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/data/ALSI.rds")

Rebalance_days <- readRDS("C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/data/Rebalance_days.rds")

#relevant packages
library(rmsfuns)
pacman::p_load("tidyr", "tbl2xts","devtools","lubridate", "readr", "PerformanceAnalytics", "ggplot2", "dplyr")
```


## Lets create an ALSI and SWIX weighted poerfolio cumulative returns 

### First lets calculate each indexes weighted daily portfolio returns usinf the safe_returns:
```{r}
#Transforming the data to xts format so we can use safe_returns
ALSI_weights_xts <- ALSI %>% 
    select(date, Tickers, J203) %>% 
    tbl_xts(cols_to_xts = J203, spread_by = Tickers)

SWIX_weights_xts <- ALSI %>% 
    select(date, Tickers, J403) %>% 
    tbl_xts(cols_to_xts = J403, spread_by = Tickers)

Returns_xts <- ALSI %>% 
    select(date, Tickers, Return) %>% 
    tbl_xts(cols_to_xts = Return, spread_by = Tickers)

#Making sure that all NA's are set to 0 

ALSI_weights_xts[is.na(ALSI_weights_xts)] <- 0
SWIX_weights_xts[is.na(SWIX_weights_xts)] <- 0
Returns_xts[is.na(Returns_xts)] <- 0

#Now using the safe returns I calculate the weighted portfolio returns for the ALSI. 
ALSI_port_returns <- rmsfuns::Safe_Return.portfolio(R = Returns_xts, weights = ALSI_weights_xts, lag_weights = T) %>%  
    xts_tbl() %>%  rename(ALSI_port_returns = portfolio.returns)

#Now using the safe returns I calculate the weighted portfolio returns for the SWIX.
SWIX_port_returns <- rmsfuns::Safe_Return.portfolio(R = Returns_xts, weights = SWIX_weights_xts, lag_weights = T) %>%  
    xts_tbl() %>%  rename(SWIX_port_returns = portfolio.returns)

#Merging the ALSI and SWIX weighted portfolio returns
returns_port_merged <- left_join(ALSI_port_returns, SWIX_port_returns, by= "date")

#Now i need to calculate the cumulative returns 

cumulative_returns_port <- returns_port_merged %>%  arrange(date) %>% 
    
    mutate(across(.cols = -date, .fns = ~cumprod(1 + .))) %>% 
    
    mutate(across(.cols = -date, .fns = ~./first(.))) %>% 
    # Start at 1
    
    pivot_longer(cols=-date, names_to = "Index", values_to = "Cummulative_port_returns")

#Now I am able to plot the cumulative portfolio returns of the ALSI and SWIX

cumulative_returns_port_plot <-    cumulative_returns_port %>%  
       
       ggplot() + 
  
  geom_line(aes(date, Cummulative_port_returns , color = Index), size = 0.5, alpha = 0.8) +
  
   fmxdat::theme_fmx(title.size = fmxdat::ggpts(25), 
                    subtitle.size = fmxdat::ggpts(0),
                    caption.size = fmxdat::ggpts(20),
                    CustomCaption = T) + 
    
  fmxdat::fmx_cols() + 
  
  labs(x = "", y = "Percentage(%)", caption = "Note:\nCalculation own",
       title = "Cumulative Returns of ALSI and SWIX Indexes",
       subtitle = "")

# Finplot for finishing touches:

fmxdat::finplot(cumulative_returns_port_plot, x.vert = T, x.date.type = "%Y", x.date.dist = "1 year")
```

As can be seen from the above graph, the cumulative returns for the ALSI and the SWIX were very similar until 2020, at which point the ALSI began outperforming the SWIX. 

#### Each sectors weighted return contribution

```{r}
portfolio_returns <- ALSI %>% 
    mutate(J203 = coalesce(J203,0)) %>% 
    mutate(J403 = coalesce(J403,0)) %>% 
    mutate(weighted_ALSI_ret = Return*J203) %>% 
    mutate(weighted_SWIX_ret = Return*J403)

portfolio_returns_with_total_sector_weights<-portfolio_returns %>% 
    arrange(date) %>% 
    group_by(date,Sector) %>% 
    mutate(ALSI_sector_weight = coalesce(J203,0)) %>% 
    mutate(SWIX_sector_weight = coalesce(J403,0)) %>% 
    mutate(ALSI_sector_returns = sum(weighted_ALSI_ret,na.rm = T)) %>% 
    mutate(total_ALSI_sector_weight = sum(J203, na.rm=T)) %>% 
    mutate(SWIX_sector_returns = sum(weighted_SWIX_ret, na.rm = T)) %>% 
    mutate(total_SWIX_sector_weight = sum(J403, na.rm=T)) %>% 
    ungroup()

#Transforming the sector weights data to xts to make use of the performance analytics package 
total_ALSI_sector_weight <-  portfolio_returns_with_total_sector_weights %>%  select(date, Sector ,total_ALSI_sector_weight) %>%  group_by(Sector) %>%  unique() %>%  
    tbl_xts(cols_to_xts = total_ALSI_sector_weight, spread_by = Sector)

total_SWIX_sector_weight <-  portfolio_returns_with_total_sector_weights %>%  select(date, Sector ,total_SWIX_sector_weight) %>%  group_by(Sector) %>%  unique() %>%  
    tbl_xts(cols_to_xts = total_SWIX_sector_weight, spread_by = Sector)

#Transforming the sectors returns to xts data so that we can make use of the performance analytics package
ALSI_sector_returns <- portfolio_returns_with_total_sector_weights %>%  select(date, Sector ,ALSI_sector_returns) %>%  group_by(Sector) %>%  unique() %>%  
    tbl_xts(cols_to_xts = ALSI_sector_returns, spread_by = Sector)

SWIX_sector_returns <- portfolio_returns_with_total_sector_weights %>%  select(date, Sector ,SWIX_sector_returns) %>%  group_by(Sector) %>%  unique() %>%  
    tbl_xts(cols_to_xts = SWIX_sector_returns, spread_by = Sector)

Fund_size_at_start <- 1000 #Setting the starting value for the portfolio 

ALSI_sector_return_division <- 
      rmsfuns::Safe_Return.portfolio(ALSI_sector_returns, 
                                     
                       weights = total_ALSI_sector_weight, lag_weights = TRUE,
                       
                       verbose = TRUE, contribution = TRUE, 
                       
                       value = Fund_size_at_start, geometric = TRUE) 

SWIX_sector_return_division <- 
      rmsfuns::Safe_Return.portfolio(SWIX_sector_returns, 
                                     
                       weights = total_SWIX_sector_weight, lag_weights = TRUE,
                       
                       verbose = TRUE, contribution = TRUE, 
                       
                       value = Fund_size_at_start, geometric = TRUE) 

```

#### The ALSI's individual sectors weighted return contribution

```{r}
#Lets now plot
 ALSI_sector_return_division$BOP.Weight  %>% .[endpoints(.,'months')] %>% chart.StackedBar()
```
#### The SWIX's individual sectors weighted return contribution

```{r}
#Lets now plot
 SWIX_sector_return_division$BOP.Weight  %>% .[endpoints(.,'months')] %>% chart.StackedBar()
```

## 5% capped portfolio

```{r}



```



# Question 4

What is needed from me in this question:
- Analyze if funds that have performed well in the past will also perform well in the future. 
- Evaluate whether funds that perform well in the past attract more investors. 
- Evaluate whether investing in past good performers is a good strategy. 


## Loading the data and relevant packages

```{r}
#Loading the data 
ASISA_Flows <- readRDS("C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/data/ASISA_Flows.rds")

ASISA_Rets <- readRDS("C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/data/ASISA_Rets.rds")

#Loading the relevant packages 
#install.packages("ggridges")
library(ggridges)
library(dplyr)
library(ggplot2)
library(tidyr)
library(lubridate)
library(knitr)
library(broom)


```

## Data preperation 

```{r}
# Merge the datasets
asisa_data <- left_join(ASISA_Flows, ASISA_Rets, by = c("date", "Fund", "Index", "FoF"))

# Ensure date columns are in Date format
asisa_data$date <- as.Date(asisa_data$date)

# Filter out Fund of Funds 
asisa_data <- filter(asisa_data, FoF == "No")

```

### Defining the look-back period and future performance periods 

```{r}
# Define look-back and future performance periods
look_back_years <- 3
future_performance_years <- 1

# Create year column for easier calculations
asisa_data <- asisa_data %>%
  mutate(year = year(date))

```

### Calculating cumulative returns for the look back period

```{r}
# Calculate cumulative returns for each fund over the look-back period
cumulative_returns <- asisa_data %>%
  group_by(Fund) %>%
  filter(year >= max(year) - look_back_years) %>%
  summarize(cumulative_return = prod(1 + Returns) - 1) %>%
  ungroup()

# Rank funds based on these returns
cumulative_returns <- cumulative_returns %>%
  arrange(desc(cumulative_return)) %>%
  mutate(rank = row_number())

```

### Analyizing fund flows and future performance

```{r}
# Merge the rankings back to the main data
asisa_data <- asisa_data %>%
  left_join(cumulative_returns, by = "Fund")

# Analyze fund flows for the year following the look-back period
fund_flows_analysis <- asisa_data %>%
  filter(year == max(year) - look_back_years + future_performance_years) %>%
  group_by(Fund) %>%
  summarize(total_flow = sum(Flows))

# Calculate future performance
future_performance <- asisa_data %>%
  filter(year >= max(year) - look_back_years + future_performance_years,
         year < max(year) - look_back_years + future_performance_years + future_performance_years) %>%
  group_by(Fund) %>%
  summarize(future_return = prod(1 + Returns) - 1) %>%
  ungroup()

```

### Analyizing data 

```{r}
# Combine flow analysis and future performance
combined_analysis <- fund_flows_analysis %>%
  left_join(future_performance, by = "Fund") %>%
  left_join(cumulative_returns, by = "Fund")



```

### Scatter plot illustrating the relationship between past performance and future performance

```{r}

# Create a new variable for coloring
combined_analysis <- combined_analysis %>%
  mutate(intensity = cumulative_return + future_return)

# Plot with color 
ggplot(combined_analysis, aes(x = cumulative_return, y = future_return, color = intensity)) +
  geom_point() +
  scale_color_gradient(low = "blue", high = "red") +  # Adjust colors as needed
  labs(title = "Past vs Future Performance of Funds",
       x = "Past Cumulative Return",
       y = "Future Return",
       color = "Intensity") +
  theme_minimal() +
  theme(legend.position = "right")


```
As can be seen from the above graph, there does seem to be a positive correlation between past cumulative returns and future performance. However, not so straight forward, as there are some funds that had a poor past cumulative return but had good future returns. Lets do some further analysis. 


### correlation analysis between cumulative past return(3 years) and total flow

```{r}
# Calculate Pearson's correlation coefficient
correlation_result <- cor.test(combined_analysis$cumulative_return, combined_analysis$total_flow, method = "pearson")

# Print the results
correlation_result

#Creating a nice table for the above results 
correlation_results_df <- data.frame(
  Statistic = c("Correlation Coefficient", "P-Value", "95% Confidence Interval Lower Bound", "95% Confidence Interval Upper Bound"),
  Value = c(0.06977338, 0.04186, 0.002581492, 0.136338075)
)


kable(correlation_results_df, 
      col.names = c("Statistic", "Value"), 
      caption = "Correlation Analysis Results", 
      align = 'c')
```

### Lets create a linear regression to see if past performance does indicate future performance. 

```{r}
# Assuming combined_analysis has 'cumulative_return' for past performance
# and 'future_return' for future performance

# Linear Regression
model <- lm(future_return ~ cumulative_return, data = combined_analysis)

# Summary of the model
summary(model)

#Making the model tidy and formating nice
# Tidy the model summary
tidy_model <- tidy(model)

# Create a table with knitr::kable
kable(tidy_model, 
      caption = "Linear Regression Model Summary", 
      align = 'c', 
      digits = 4)

```
This regression analysis shows that there is a statistically significant relationship between past return values and future return values. Thus funds with higher past returns tend to have higher future returns. However the R-squared value is fairly low at 0.12 which suggests that past performance only explain 12% of the variation in future performance. Thus although past performance does have a statistically significant positive relationship with future performance, this relationship is fairly week and should not be soley relied upon. 


# Question 5 




# Question 6 


