---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Question 3: Portfolio construction and comparison"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Austin Byrne"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Stellenbosch University, Stellenbosch, South Africa" # First Author's Affiliation
Email1: "22582053\\@sun.ac.za" # First Author's Email address

#Author2: "John Smith"
#Ref2: "Some other Institution, Cape Town, South Africa"
#Email2: "John\\@gmail.com"
#CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.

#Author3: "John Doe"
#Email3: "Joe\\@gmail.com"

CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
#keywords: "Portfolio analysis \\ALSI \\SWIX" # Use \\sep to separate
#JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
abstract: |
  The purpose of this question is to construct a portfolio analysis on the weighted ALSI and weighted SWIX portfolios. In this question I will plot the cumulative returns of both portfolios and evaluate the comparisons and differences. Furthermore, a breakdown of the various weighted contributions of each sector within each weighted portfolio will be analysed. 
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
# Note: Include = FALSE implies the code is executed, but not printed in your pdf.
# warning and message = FALSE implies ugly messages and warnings are removed from your pdf.
# These should be picked up when you execute the command chunks (code sections below) in your rmd, not printed in your paper!

# Lets load in example data, and see how this can be stored and later called from your 'data' folder.
if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
Example_data <- Texevier::Ex_Dat

# Notice that as you are working in a .Rproj file (I am assuming you are) - the relative paths of your directories start at your specified root.
# This means that when working in a .Rproj file, you never need to use getwd() - it is assumed as your base root automatically.
write_rds(Example_data, path = "data/Example_data.rds")

```


<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Introduction \label{Introduction}

Using the information on the ALSI (J203) and SWIX (J403) Indexes, I will be writing a short report where I will be comparing the specific methodologies used by the ALSI and the SWIX, evaluating the cumulative returns of each and plotting them together to perform a comparison, evaluating which sectors hold the highest weight and contribution for each. 

# Loading relevant data and packages 

The data used is data based on the ALSI and SWIX. The data contains dates, weights and returns for each. 

```{r}
#Relevant data
ALSI <- readRDS("C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/data/ALSI.rds")

Rebalance_days <- readRDS("C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/data/Rebalance_days.rds")

#relevant packages
library(rmsfuns)
pacman::p_load("tidyr", "tbl2xts","devtools","lubridate", "readr", "PerformanceAnalytics", "ggplot2", "dplyr")
```

# Lets create an ALSI and SWIX weighted poerfolio cumulative returns 

## First lets calculate each indexes weighted daily portfolio returns using the safe_returns:

In order to use the safe_returns function to calculate the weighted portfolio returns I need to convert the data to and xts format. That is done below: 

```{r}
#Transforming the data to xts format so we can use safe_returns
ALSI_weights_xts <- ALSI %>% 
    select(date, Tickers, J203) %>% 
    tbl_xts(cols_to_xts = J203, spread_by = Tickers)

SWIX_weights_xts <- ALSI %>% 
    select(date, Tickers, J403) %>% 
    tbl_xts(cols_to_xts = J403, spread_by = Tickers)

Returns_xts <- ALSI %>% 
    select(date, Tickers, Return) %>% 
    tbl_xts(cols_to_xts = Return, spread_by = Tickers)

#Making sure that all NA's are set to 0 

ALSI_weights_xts[is.na(ALSI_weights_xts)] <- 0
SWIX_weights_xts[is.na(SWIX_weights_xts)] <- 0
Returns_xts[is.na(Returns_xts)] <- 0

#Now using the safe returns I calculate the weighted portfolio returns for the ALSI. 
ALSI_port_returns <- rmsfuns::Safe_Return.portfolio(R = Returns_xts, weights = ALSI_weights_xts, lag_weights = T) %>%  
    xts_tbl() %>%  rename(ALSI_port_returns = portfolio.returns)

#Now using the safe returns I calculate the weighted portfolio returns for the SWIX.
SWIX_port_returns <- rmsfuns::Safe_Return.portfolio(R = Returns_xts, weights = SWIX_weights_xts, lag_weights = T) %>%  
    xts_tbl() %>%  rename(SWIX_port_returns = portfolio.returns)

#Merging the ALSI and SWIX weighted portfolio returns
returns_port_merged <- left_join(ALSI_port_returns, SWIX_port_returns, by= "date")
#We now have a merged data frame that contains the returns for the weighted ALSI and SWIX portfolios. From here we can do some cool analysis. Which will be done now. 
```

Now that we have the merged data frame of portfolio returns for the ALSI and SWIX in the correct format, I can move onto calculating the cumulative weighted returns. 

```{r}
#Now i need to calculate the cumulative returns. Calculating the cumulative returns for each weighted portfolio will  provide a good base to do meaningful comparison.  


cumulative_returns_port <- returns_port_merged %>%  arrange(date) %>% 
    
    mutate(across(.cols = -date, .fns = ~cumprod(1 + .))) %>% 
    
    mutate(across(.cols = -date, .fns = ~./first(.))) %>% 
    # Start at 1
    
    pivot_longer(cols=-date, names_to = "Index", values_to = "Cummulative_port_returns")



cumulative_returns_port_plot <-    cumulative_returns_port %>%  
       
       ggplot() + 
  
  geom_line(aes(date, Cummulative_port_returns , color = Index), size = 0.5, alpha = 0.8) +
  
   fmxdat::theme_fmx(title.size = fmxdat::ggpts(25), 
                    subtitle.size = fmxdat::ggpts(0),
                    caption.size = fmxdat::ggpts(20),
                    CustomCaption = T) + 
    
  fmxdat::fmx_cols() + 
  
  labs(x = "", y = "Percentage(%)", caption = "Note:\nCalculation own",
       title = "Cumulative Returns of ALSI and SWIX Indexes",
       subtitle = "")
```

## Cumulative returns plot

```{r Figure 1, warning =  FALSE, fig.align = 'center', fig.cap = "Cumulative returns plot \\label{Figure1}", fig.height = 6, fig.width = 6, dev = 'png'}
# Finplot for finishing touches:

fmxdat::finplot(cumulative_returns_port_plot, x.vert = T, x.date.type = "%Y", x.date.dist = "1 year")
```

As can be seen from the above graph, the cumulative returns for the ALSI and the SWIX were very similar until 2020, at which point the ALSI began outperforming the SWIX. Immediate prior to this separation in cumulative returns was a huge draw down for both the ALSI and SWIX weighted portfolios. Since this draw down is within 2020, a possible reason could be the COVID-19 pandemic. There may have been a switch in weights between the two portfolios which would result in varying returns. From the graph it is evident that the ALSI reacted better than the SWIX to this heavy draw down in 2020. 

To further this analysis I will dive into the weighted construction of the ALSI and SWIX portfolios, more precisely, I will be analyzing the weighted return contribution for each sector and evaluating the differences in the ALSI and SWIX. 

## Each sectors weighted return contribution

In this section of code I am grouping by sector to then be able to calculate the weighted reuturn contribution of each sector in the respective ALSI and SWIX portfolios. I can then evaluate which sectors are the main drivers to portfolio returns.

```{r}
#Creating a data frame that contains weighted portfolio returns for the ALSI and SWIX
portfolio_returns <- ALSI %>% 
    mutate(J203 = coalesce(J203,0)) %>% 
    mutate(J403 = coalesce(J403,0)) %>% 
    mutate(weighted_ALSI_ret = Return*J203) %>% 
    mutate(weighted_SWIX_ret = Return*J403)

#In this section of code I am grouping by sector to then be able to calculate the weighted reuturn contribution of each sector in the respective ALSI and SWIX portfolios. I can then evaluate which sectors are the main drivers to portfolio returns. 
portfolio_returns_with_total_sector_weights<-portfolio_returns %>% 
    arrange(date) %>% 
    group_by(date,Sector) %>% 
    mutate(ALSI_sector_weight = coalesce(J203,0)) %>% 
    mutate(SWIX_sector_weight = coalesce(J403,0)) %>% 
    mutate(ALSI_sector_returns = sum(weighted_ALSI_ret,na.rm = T)) %>% 
    mutate(total_ALSI_sector_weight = sum(J203, na.rm=T)) %>% 
    mutate(SWIX_sector_returns = sum(weighted_SWIX_ret, na.rm = T)) %>% 
    mutate(total_SWIX_sector_weight = sum(J403, na.rm=T)) %>% 
    ungroup()

#Transforming the sector weights data to xts to make use of the performance analytics package 
total_ALSI_sector_weight <-  portfolio_returns_with_total_sector_weights %>%  select(date, Sector ,total_ALSI_sector_weight) %>%  group_by(Sector) %>%  unique() %>%  
    tbl_xts(cols_to_xts = total_ALSI_sector_weight, spread_by = Sector)

total_SWIX_sector_weight <-  portfolio_returns_with_total_sector_weights %>%  select(date, Sector ,total_SWIX_sector_weight) %>%  group_by(Sector) %>%  unique() %>%  
    tbl_xts(cols_to_xts = total_SWIX_sector_weight, spread_by = Sector)

#Transforming the sectors returns to xts data so that we can make use of the performance analytics package
ALSI_sector_returns <- portfolio_returns_with_total_sector_weights %>%  select(date, Sector ,ALSI_sector_returns) %>%  group_by(Sector) %>%  unique() %>%  
    tbl_xts(cols_to_xts = ALSI_sector_returns, spread_by = Sector)

SWIX_sector_returns <- portfolio_returns_with_total_sector_weights %>%  select(date, Sector ,SWIX_sector_returns) %>%  group_by(Sector) %>%  unique() %>%  
    tbl_xts(cols_to_xts = SWIX_sector_returns, spread_by = Sector)

#Lets make an arbitrary starting amount for our funds. 
Fund_size_at_start <- 1000 #Setting the starting value for the portfolio 

ALSI_sector_return_division <- 
      rmsfuns::Safe_Return.portfolio(ALSI_sector_returns, 
                                     
                       weights = total_ALSI_sector_weight, lag_weights = TRUE,
                       
                       verbose = TRUE, contribution = TRUE, 
                       
                       value = Fund_size_at_start, geometric = TRUE) 

SWIX_sector_return_division <- 
      rmsfuns::Safe_Return.portfolio(SWIX_sector_returns, 
                                     
                       weights = total_SWIX_sector_weight, lag_weights = TRUE,
                       
                       verbose = TRUE, contribution = TRUE, 
                       
                       value = Fund_size_at_start, geometric = TRUE) 

```

## The ALSI's individual sectors weighted return contribution

Here we are able to evaluate which sector(s) are the main weighted return drivers of the ALSI weighted portfolio. From the plot below it is evident that for the last 10 years the majority of the return in the ALSI weighted portfolio has come from the Industrial sector and the lowest coming from the property/residual sectors. However, since 2016 the industrial sector has slowly started lossing its majority hold to an ever rising contribution from the resources sector. 

Next we are going to evaluate the same plot but for the SWIX weighted portfolio. We will then be able to do identify a possibility of why the cumulative returns of the two portfolios differ. 

```{r Figure 2, warning =  FALSE, fig.align = 'center', fig.cap = "ALSI sector contribution plot \\label{Figure2}", fig.height = 6, fig.width = 6, dev = 'png'}
#Lets now plot
 ALSI_sector_return_division$BOP.Weight  %>% .[endpoints(.,'months')] %>% chart.StackedBar()
```

## The SWIX's individual sectors weighted return contribution

The plot below evaluates which sector(s) are the main weighted return drivers of the SWIX weighted portfolio. Like that of the ALSI, the majority holding is found in the industrial sector with again the minority held by the the property/residual sectors. The trajectory of increased contribution from the resources sector is scene here again as was seen in the ALSI portfolio. However, the differences lie in the actual value. It is evident that the SWIX weighted portfolio has a more even contribution between the financial sector and resources sector. Where the ALSI has a higher contribution from the resources sector over the financial sector.  

```{r Figure 3, warning =  FALSE, fig.align = 'center', fig.cap = "SWIX sector contribution plot \\label{Figure3}", fig.height = 6, fig.width = 6, dev = 'png'}

#Lets now plot
 SWIX_sector_return_division$BOP.Weight  %>% .[endpoints(.,'months')] %>% chart.StackedBar()
```

## Comparisson of the ALSI and SWIX portfolio contribution

When evaluating the weighted contributions of the various sectors to the ALSI and SWIX portfolios, there are some similarities but also some points of difference. Both portfolios have a large contribution from the industrial sector. Where the differences play a part is with respect to the financial a resources sectors. The ALSI obtains more weight from the resource sectors over that of the financial sector, while the SWIX obtains more of an even split between the two. This difference may be the contributing factor as to why the ALSI performed better than the SWIX since 2020. 

<!-- Make title of bibliography here: -->
<!-- \newpage -->

\newpage

# References {-}

<div id="refs"></div>


# Appendix {-}

## Appendix A {-}

Some appendix information here

## Appendix B {-}

