---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Question 5: Volatility and GARCH estimates"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Austin Byrne"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Stellenbosch University, Western Cape" # First Author's Affiliation
Email1: "22582053\\@sun.ac.za" # First Author's Email address

#Author2: "John Smith"
#Ref2: "Some other Institution, Cape Town, South Africa"
#Email2: "John\\@gmail.com"
#CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.

#Author3: "John Doe"
#Email3: "Joe\\@gmail.com"

#CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
#keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
#JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
#BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
#addtoprule: TRUE
#addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
abstract: |
  This question is broken down into two sections, In section 1, I evaluate the statement that the rand has been one of the most volatile currencies in recent years. Tackling this question I calculate the standard deviation of nine currencies and compare. From the plots it is evident that the Rand is one of the most volatile in recent years. In the second section I attampt to tackle the problem using a GARCH model. 
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
# Note: Include = FALSE implies the code is executed, but not printed in your pdf.
# warning and message = FALSE implies ugly messages and warnings are removed from your pdf.
# These should be picked up when you execute the command chunks (code sections below) in your rmd, not printed in your paper!

# Lets load in example data, and see how this can be stored and later called from your 'data' folder.
if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)

```


<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Introduction \label{Introduction}

In the first part of the question I analyze the statement that the Rand has been one of the most volatile stocks in recent years. For my analysis I find eight other currencies of which I can compare the volatility of the Rand to. The choice of which currencies to compare is base on historical performance. I first analyse the standard deviation of the currencies from prior 1990 to 2021. Over this long period it is not evident that the rand is one of the mos volatile. More specifically the questions states over recent years. Thus fr the next part of the analysis I shorten the time horizon to three years, comparing the volatility of the currencies in question to that of the Rand from 2018 to 2021. Here it is more evident that the Rand is one of the most volatile currencies. However, the Rand is not drastically more volatile than the Singapore, Brazilian and Indian currencies. 

In the next section I dive into a GARCH model analysis. Unfortunetly I ran out of time to do this part of the question  

## Loading relvant data from the question

```{r}
cncy <- readRDS("C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/data/currencies.rds")
cncy_carry <- readRDS("C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/data/cncy_Carry.rds")
cncy_value <- readRDS("C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/data/cncy_value.rds")
cncyIV <- readRDS("C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/data/cncyIV.rds")
bbdxy <- readRDS("C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/data/bbdxy.rds")
```

## Evaluating the volatility of the Rand against other currencies

In the first part of this question I need to comment on the statement that the Rand has been one of the most volatile currencies. To start this analysis I will first prepare the data available.

### Data preperation 

For this section of the analysis I will be using the cncy data set which contains information on the currency values of 41 currencies. For the analysis it will an over whelming amount of information if I compare the ZAR to all of these currencies. I will therefore pick a few to do the comparison on. The way in which I will choose which currencies to compare the rand to is based of prior knowledge. I want to compare the rand to the historically most volatile currencies as well as the historically least volatile currencies and currencies that are historically moderately volatile. 

For the historically most volatile currencies I will be evaluating the Argentinian currency, the Turkish currency and the Brazilian currency. For the historically least volatile currencies I will be evaluating Singapore currency, the Canadian currency and the Australian currency. For the historically moderate currencies I will be evaluating Mexico and India. 

```{r}
#Collecting the currency that I want to evaluate into one vector
comparison_countries <- c("Argentina_Cncy","Turkey_Cncy","Brazil_Cncy","Singapore_Cncy","Canada_Cncy","Australia_Cncy_Inv","Mexico_Cncy","India_Cncy", "SouthAfrica_Cncy")

#Now lets filter our data to have a data frame with just these countries that I want to consider
comparison_cncy <- cncy %>%  filter(Name %in% comparison_countries) 

#Now lets clean up this data by removing the common "_cncy". 
comparison_cncy <- comparison_cncy %>% 
    mutate(Name = gsub( "_Cncy", "", Name)) 

```


## Volatility comparison plot from prior 1993 for 9 countries

From the plot below it is not plainly evident that the South African currency is the most volatile. We are however, looking at a very large scale and the question pertains to recent years. Therefore to gain a better understanding of the volatility of the South African rand in recent years I will look at the last 3 years of standard deviation for each currency in question. 

```{r Figure 1, warning =  FALSE, fig.align = 'center', fig.cap = "Long term volatility comparison plot \\label{Figure1}", fig.height = 6, fig.width = 6, dev = 'png'}
comparison_cncy <- comparison_cncy %>% 
    arrange(date) %>% 
    group_by(Name) %>% 
    mutate(change = Price - lag(Price)) %>% 
    filter(date > first(date)) %>%
    mutate(SD = sqrt((change-mean(change))^2)) %>% 
    ungroup()



comparison_cncy_SD_plot <- comparison_cncy %>%   

  ggplot() + 
  
   geom_line(aes(date, SD , color = Name), size = 0.5, alpha = 0.8) +
    
    facet_wrap(~Name, scales = "free_y")+
  
   fmxdat::theme_fmx(title.size = fmxdat::ggpts(30), 
                    subtitle.size = fmxdat::ggpts(0),
                    caption.size = fmxdat::ggpts(25),
                    CustomCaption = T) + 
    
  fmxdat::fmx_cols() + 
  
  labs(x = "date", y = "%", caption = "Note:\nCalculation own",
       title = "Standard deviation comparison plot",
       subtitle = "")

# Finplot for finishing touches:
fmxdat::finplot(comparison_cncy_SD_plot, x.vert = T, x.date.type = "%Y", x.date.dist = "5 years", darkcol = F)

```

## Volatility comparison plot over last 3 years for 9 countries

To gain a better understanding of the volatility of the Rand relative to its peers in the last few years I will be shorten the time horison inspected. The data ends in 2021. I will therefore look at the last 3 years of standard deviation. More specifically, from 2018.  

From the more recent volatility plot comparison, the Rand does seem to be one of the more volatile currencies in our data set. However, it is not strikingly more volatile than Singapore, India or Brazil. Thus, to conclude, the Rand has been on eo fhte more volatile currencies in recent years. 

```{r Figure 2, warning =  FALSE, fig.align = 'center', fig.cap = "3 year volatility comparison plot \\label{Figure2}", fig.height = 6, fig.width = 6, dev = 'png'}
shortened_comparison_cncy <- comparison_cncy %>% 
    filter(date > lubridate::ymd(20180101))

shortened_comparison_cncy_SD_plot <- shortened_comparison_cncy %>%   

  ggplot() + 
  
   geom_line(aes(date, SD , color = Name), size = 0.5, alpha = 0.8) +
    
    facet_wrap(~Name, scales = "free_y")+
  
   fmxdat::theme_fmx(title.size = fmxdat::ggpts(30), 
                    subtitle.size = fmxdat::ggpts(0),
                    caption.size = fmxdat::ggpts(25),
                    CustomCaption = T) + 
    
  fmxdat::fmx_cols() + 
  
  labs(x = "date", y = "%", caption = "Note:\nCalculation own",
       title = "Standard deviation comparison plot",
       subtitle = "")

# Finplot for finishing touches:
fmxdat::finplot(shortened_comparison_cncy_SD_plot, x.vert = T, x.date.type = "%Y", x.date.dist = "1 years", darkcol = F)

```



# References {-}

<div id="refs"></div>


# Appendix {-}

## Appendix A {-}

Some appendix information here

## Appendix B {-}

