---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Question 6: Portfolio Construction"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Austin Byrne"  # First Author - note the thanks message displayed as an italic footnote of first page.
#Ref1: "Some Institution, Cape Town, South Africa" # First Author's Affiliation
#Email1: "nfkatzke\\@gmail.com" # First Author's Email address

#Author2: "John Smith"
#Ref2: "Some other Institution, Cape Town, South Africa"
#Email2: "John\\@gmail.com"
#CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.

#Author3: "John Doe"
#Email3: "Joe\\@gmail.com"

#CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
#keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
#JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
abstract: |
  In this question I create optimal portfolios making use of Minimum Variance Optimization. I first construct a unconstrained portfolio and then compare it to the constrained portfolio. The constrained portfolio results in lower risk and lower return.  
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
# Note: Include = FALSE implies the code is executed, but not printed in your pdf.
# warning and message = FALSE implies ugly messages and warnings are removed from your pdf.
# These should be picked up when you execute the command chunks (code sections below) in your rmd, not printed in your paper!

# Lets load in example data, and see how this can be stored and later called from your 'data' folder.
if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)


```


<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Introduction \label{Introduction}

In this question I dive into the realm of portfolio construction. To create an optimal portfolio I make use of the Minimum Variance Optimizer (MVO) which creates an optimal portfolio with the lowest possible variance. I construct an unconstrained and constrained portfolio using MVO. The resulting optimal portfolios are low risk low return portfolios which can be expected from using MVO. The constrained optimal portfolio results in lower risk and lower annualized monthly returns.

## Laod relevant functions

I load any function that I may have created so that I can use them in this code and analysis. 

```{r}
library(purrr)
list.files('C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/code', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))
```


## Loading relevant data and packages

Here I load the relevant data. 

```{r}
MAA <- readRDS("C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/data/MAA.rds")

msci <- readRDS("C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/data/msci.rds")  %>% filter(Name %in% c("MSCI_ACWI", "MSCI_USA", "MSCI_RE", "MSCI_Jap"))

```


## Data preperation: Lets convert the price data to monthly returns data

In this section I do some data perpetration. I take the original data of the MAA and msci and create a new column called Monthly_returns, which calculates the monthly returns from the daily data. I also filter for post 2010 and ensure that at least three years a data is available.

### MAA monthly returns 

Monthly returns for MAA are calculated after some fun wrangling.

```{r}
# Adding Year and Month columns without disturbing the original data
MAA <- MAA %>%
  mutate(
    Year = year(date),
    Month = month(date)
  )

# Calculating monthly returns..
MAA <- MAA %>%
  group_by(Ticker, Year, Month) %>%
  arrange(date, .by_group = TRUE) %>%
  summarize(
    Monthly_Return = last(Price) / first(Price) - 1,
    .groups = 'drop'
  ) %>%
  # Joining the calculated monthly returns with the original data
  left_join(MAA, by = c("Ticker", "Year", "Month"))

MAA <- MAA %>% 
    filter(date > as.Date('2010-01-01')) %>% 
    filter(date >= min(date) + years(3))
```

### msci monthly returns

Monthly returns for msci are calculated after some fun wrangling.

```{r}
msci <- msci %>%
  mutate(Year = year(date),Month = month(date))

# Calculating monthly returns..
msci <- msci %>%
  group_by(Name, Year, Month) %>%
  arrange(date, .by_group = TRUE) %>%
  summarize(
    Monthly_Return = last(Price) / first(Price) - 1,
    .groups = 'drop'
  ) %>%
  # Joining the calculated monthly returns with the original data
  left_join(msci, by = c("Name", "Year", "Month"))

msci <- msci %>% 
    filter(date > as.Date('2010-01-01')) %>% 
    filter(date >= min(date) + years(3))
```

### Creating an asset class column in my data sets

There are now four different asset classes in the MAA data set, namely, "Assian currency", "Credit", "Rates", "Commodity", "Equity", and "Real estate.

```{r}
#Using created function to add asset class to data sets
MAA <- add_asset_class(MAA, asset_classes)
msci <- add_asset_class(msci, asset_classes)
```

## Lets now attempt to construct the portfolio by making use of mean variance optimization: 

I now make use of  ean VAriance Optimization to first create a unconstrained portfolio. I am then able to evaluate the risk and return of the unconstrained portfolio which I then compare to the constrained portfolio. 

### Combine the Data Sets and compute some graphs

Here I make use of a function I created that prepares and combines the relavant data. 

```{r}
combined_data <- prepare_and_combine(MAA, msci)
```

### Rolling 3 year annualzied returns comparison of asset classes

TO insoect the data in question I run a plot that conatins the 3 year annualsied returns for each asset class. What is immidatley evident in this plot is that the equities asset class is the most volatile which also means the asset class that potentially holds the most risk. Outside of the equities asset class, the remaining asset classes are relatively similiar, with USD beinf the next most volatile. 

```{r}
#Calculating rolling return on a three year annualized basis
combined_data <- combined_data %>% arrange(date) %>% 
    group_by(Asset_Class) %>% 
    mutate(Roll_rets = RcppRoll::roll_prod(1+ Monthly_Return, 36, fill = NA, align = "right")^(12/36) -1) %>%
    group_by(date) %>%  
    filter(any(!is.na(Roll_rets))) %>%  # Remove dates with no rolling returns
    ungroup()

#Now lets create the graph
Asset_rolling_Comparisson_plot <- combined_data %>%    
    
    ggplot() + 
  
  geom_line(aes(date, Roll_rets, color = Asset_Class), size = 0.9, alpha = 0.7) + 
  
  fmxdat::theme_fmx(title.size = fmxdat::ggpts(30), 
                    subtitle.size = fmxdat::ggpts(25),
                    caption.size = fmxdat::ggpts(25),
                    CustomCaption = T) + 
  
  fmxdat::fmx_cols() + 
  
  labs(x = "", y = "%", caption = "Note:\nCalculation own",
       title = "Asset class rolling three year returns",
       subtitle = "High volatility in equity in recent years")


# Finplot for finishing touches easily:

  fmxdat::finplot(Asset_rolling_Comparisson_plot, x.vert = T, x.date.type = "%Y", x.date.dist = "1 years", )
```

### Boxplot of monthly returns by Asset class

Here I make use of a function I create to obtain the boxplot for each asset class which illustrates the distribution of monthly returns for each asset class. The equity asset class as expected has the highest average monthly return but also contains alot of outliers which can be expected after seeing how volatile the returns of equity can be. The most stable asset classes seem to be the Asian currency and USD. This plot creates a nice illustration of the spread of monthly returns for each asset class. 

```{r}
Asset_Boxplot_Comparison <- create_asset_boxplot(combined_data)
print(Asset_Boxplot_Comparison)

```


### Computing the returns matrix

In order to using the Minimum Variance Optimizer I need to first craete a returns matrix, whicg contains the returns of each asset class in a matrix. TO do this I make use of a function that was crated in the practical. 

```{r}
#Calculating mean returns and covariance matrix
# Reshape the data for covariance calculation
return_mat_Quick <- combined_data %>%
  select(date, Ticker, Monthly_Return) %>%
  spread(key = Ticker, value = Monthly_Return)


return_mat <- impute_missing_returns(return_mat_Quick, impute_returns_method = "NONE")
```


### Computing the simple covariance matrix

Like that of the returns matrix one needs to create a simple covariance matrix to make use of MVO and be able to create an optimal portfolio. 

```{r}
# we need to drop the date columns for this
return_mat_Nodate <- data.matrix(return_mat[, -1])

# Simple Sample covariance and mean:
Sigma <- RiskPortfolios::covEstimation(return_mat_Nodate)

mu <- return_mat %>% summarise(across(-date, ~prod(1+.)^(1/n())-1)) %>% purrr::as_vector()

```


### Using Minimum variance optimization (MVO) for an unrestricted portfolio.

Now using MVO I am able to construct an unconstrained optimal portfolio and evaluate the risk and returns. The risk will be evaluated by the standard deviation. 

As can be seen from the table below, The optimal unconstrained portfolio has a standard deviation of 0.09827125 and an average annualized monthly retun of 0.02648242. Thus the portfolio has a very low standard deviation which is expected since the portfolio was constructed using MVO. Since the portfolio takes on low risk, it is not suprising to see that the return is also fairly low at 0.02648242. Thus, the unconstrained optimal portfolio using MVO is a low risk low return portfolio. 

Lets now add the constraints in and then compare to this base portfolio. 


```{r}
mu_ann <- (1+mu)^(12)-1
Sigma_ann <- Sigma * sqrt(12)
N <- ncol(Sigma_ann)
Imat <- rep(1, N)
sigma_inv <- solve(Sigma_ann)

mvp_weights_raw <- sigma_inv %*% Imat
mvp_weights <- as.vector(mvp_weights_raw / sum(mvp_weights_raw))
mvp_weights_df <- tibble(Tickers = rownames(mvp_weights_raw), minv_weights = mvp_weights)

# The annualised return and SD for this portfolio follows:
tibble(
  average_ret_Ann = as.numeric(t(mvp_weights) %*% mu_ann),
  volatility_Ann = as.numeric(sqrt(t(mvp_weights) %*% Sigma_ann %*% mvp_weights))*sqrt(252)
)
```

### Now lets create the constrained optimised portfolio 

With the specific constraints being:
Apply Quarterly Rebalancing;
- Limit exposure to Bonds and credit instruments at 25%;
- Limit exposure to Equities at 60%;
- Limit single asset exposure at 40%


#### Lets first create some bond/credit and equity idicies that we can use for our constraints later

Here I create the constraints that will be used to construct the constrained optimal portfolio. 


```{r}

#Creating the equity indices constraint 
# Create a vector of unique asset tickers
unique_assets <- unique(combined_data$Ticker)

# Initialize a logical vector for equity indices
equity_indices <- rep(FALSE, length(unique_assets))

# Set TRUE for assets that are equities
equity_assets <- combined_data %>%
                  filter(Asset_Class == "Equity") %>%
                  pull(Ticker) %>%
                  unique()
equity_indices[unique_assets %in% equity_assets] <- TRUE


#Creating the bond_credit indices constraint
# Initialize a logical vector for bond and credit indices
bond_credit_indices <- rep(FALSE, length(unique_assets))

# Set TRUE for assets that are bonds or credit instruments
bond_credit_assets <- combined_data %>%
                      filter(Asset_Class %in% c("Bonds", "Credit")) %>%
                      pull(Ticker) %>%
                      unique()
bond_credit_indices[unique_assets %in% bond_credit_assets] <- TRUE




```

#Adding constraints to the optimisation problem and solving

Solving for the constrained optimal portfolio is done using the CVXR package in R. 

Lets now evaluate the constrained portfolio. The constrained optimal portfolio is an even lower risk and lower retun portfolio than that of the unconstrained portfolio created previously. The standard deviation is now at 0.0003147487 and the monthly annualised return is at 0.0003147487. Thus by adding the constrains to the portfolio we decrease the risk in the portfolio through decreased standard deviation, however, with decreased risk come decreased returns. Once again since we make use of MVO to create the optimal portfolio, it is not surprising that the optimal portfolio is a low risk low return optimal portfolio.


```{r}
pacman::p_load(knitr)
pacman::p_load(CVXR)

# Number of stocks
NStox <- ncol(return_mat_Nodate)

w <- Variable(NStox)
ret <- t(mu) %*% w
risk <- quad_form(w, Sigma)

obj <- ret - risk

# Constraints
LB = 0.01  # Minimum weight
UB = 0.40  # Maximum weight for a single asset
equity_limit = 0.60  # Maximum weight for equity assets
bond_credit_limit = 0.25 #maximum weight for bond/credit 

# Constraint list
constr <- list(w >= LB, w <= UB, sum(w) == 1, sum(w[equity_indices]) <= equity_limit, sum(w[equity_indices]) <= bond_credit_limit)

# Optimization problem
prob <- Problem(Maximize(obj), constr)

# Solve the problem
result <- solve(prob)

# Extract the optimized weights
optimized_weights <- result$getValue(w)
optimized_weights

result.CVXR <- tibble(stocks = unique_assets, weight = result$getValue(w) %>% as.vector())
print(result.CVXR)

result$getValue(risk)
result$getValue(ret)


# Create a data frame with the results
results_df <- data.frame(
  Metric = c("Annualized Return", "Annualized Risk"),
  Value = c(result$getValue(ret), result$getValue(risk))
)

# Format the table
kable(results_df, col.names = c("Metric", "Value"), caption = "Portfolio Performance Metrics")

```



\newpage

# References {-}

<div id="refs"></div>


# Appendix {-}

## Appendix A {-}

Some appendix information here

## Appendix B {-}

