---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Question 4: Flows analysis"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Austin Byrne"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "The best quantitative analyst ever seen, Stellenbosh" # First Author's Affiliation
Email1: "22582053\\@sun.ac.za" # First Author's Email address

#Author2: "John Smith"
#Ref2: "Some other Institution, Cape Town, South Africa"
#Email2: "John\\@gmail.com"
#CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.

#Author3: "John Doe"
#Email3: "Joe\\@gmail.com"

#CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
#keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
#JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
abstract: |
  In this question I evaluate whether past performance of a fund is a good indicator of future performance and whther this pushes active investors to invest in the high performing funds solely on past  performance. In this question I evaluate two look back periods, one on the shorter spectrum being 3 years and another on the longer being 10 years. I draw up scatter plots and run correlation and regression analysis to perform my analysis. 

---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
# Note: Include = FALSE implies the code is executed, but not printed in your pdf.
# warning and message = FALSE implies ugly messages and warnings are removed from your pdf.
# These should be picked up when you execute the command chunks (code sections below) in your rmd, not printed in your paper!

# Lets load in example data, and see how this can be stored and later called from your 'data' folder.
if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)

```


<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Introduction \label{Introduction}

This question will provide insight into industry fund flows between actively managed funds. More specifically, I will firstly be evaluating whether funds that have performed well in the past provides an indicator to investors that they will perform well in the future. I will then be evaluating whether this is a good investment strategy, to continually invest in funds that have performed well in the past. A crucial component of this analysis is the look back period, ie) how long must a fund perform well before it is indacted as an above average performing fund. In this question I will be evaluating a look back period of 3 years and 10 years. 

## Loading the data and relevant packages

The relevant data that is loaded here contains information on the various actively traded funds, along with trading flow data and returns. 

```{r}
#Loading the data 
ASISA_Flows <- readRDS("C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/data/ASISA_Flows.rds")

ASISA_Rets <- readRDS("C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/data/ASISA_Rets.rds")

#Loading the relevant packages 
#install.packages("ggridges")
library(ggridges)
library(dplyr)
library(ggplot2)
library(tidyr)
library(lubridate)
library(knitr)
library(broom)


```

## Data preperation 

To prepare this data for analysis I join the two flows and returns data sets, insure that the date is of the correct format and filter out any "Fund of Funds" data. 

```{r}
# Merge the datasets
asisa_data <- left_join(ASISA_Flows, ASISA_Rets, by = c("date", "Fund", "Index", "FoF"))

# Ensure date columns are in Date format
asisa_data$date <- as.Date(asisa_data$date)

# Filter out Fund of Funds 
asisa_data <- filter(asisa_data, FoF == "No")

```

## Defining the look-back period and future performance periods 

For the first comparison I will be evaluating a look back period of three years. This look back period is the shortest of the look back periods that will be evaluated. The main takeaway from this analysis is as follows; if a fund performs well for a fairly short period of three years, does this indicate that the future is bright for the fund and will this attract new investment? 

```{r}
# Define look-back and future performance periods
look_back_years <- 3
future_performance_years <- 1

# Create year column for easier calculations
asisa_data <- asisa_data %>%
  mutate(year = year(date))

```

## Calculating cumulative returns for the 3 year look back period

The cumulative returns of the 3 year period is calculated so that a comparative analysis can be done. 

```{r}
# Calculate cumulative returns for each fund over the look-back period
cumulative_returns <- asisa_data %>%
  group_by(Fund) %>%
  filter(year >= max(year) - look_back_years) %>%
  summarize(cumulative_return = prod(1 + Returns) - 1) %>%
  ungroup()

# Rank funds based on these returns
cumulative_returns <- cumulative_returns %>%
  arrange(desc(cumulative_return)) %>%
  mutate(rank = row_number())

```

## Analyizing fund flows and future performance

Calculating the future performance (the performance a year after the look back period) will provide somewhat of an answer to whether the look back period was a good indicating of the following period. 

```{r}
# Merge the rankings back to the main data
asisa_data <- asisa_data %>%
  left_join(cumulative_returns, by = "Fund")

# Analyze fund flows for the year following the look-back period
fund_flows_analysis <- asisa_data %>%
  filter(year == max(year) - look_back_years + future_performance_years) %>%
  group_by(Fund) %>%
  summarize(total_flow = sum(Flows))

# Calculate future performance
future_performance <- asisa_data %>%
  filter(year >= max(year) - look_back_years + future_performance_years,
         year < max(year) - look_back_years + future_performance_years + future_performance_years) %>%
  group_by(Fund) %>%
  summarize(future_return = prod(1 + Returns) - 1) %>%
  ungroup()

```


## Analyizing data 

To analyze the data, a scatter plot, correlation analysis and regression analysis will be conducted. 

```{r}
# Combine flow analysis and future performance
combined_analysis <- fund_flows_analysis %>%
  left_join(future_performance, by = "Fund") %>%
  left_join(cumulative_returns, by = "Fund")



```

### Scatter plot illustrating the relationship between past performance and future performance

```{r Figure 1, warning =  FALSE, fig.align = 'center', fig.cap = "3 year look back scatter plot \\label{Figure1}", fig.height = 6, fig.width = 6, dev = 'png'}

# Create a new variable for coloring
combined_analysis <- combined_analysis %>%
  mutate(intensity = cumulative_return + future_return)

# Plot with color 
ggplot(combined_analysis, aes(x = cumulative_return, y = future_return, color = intensity)) +
  geom_point() +
  scale_color_gradient(low = "blue", high = "red") +  # Adjust colors as needed
  labs(title = "Past vs Future Performance of Funds",
       x = "Past Cumulative Return",
       y = "Future Return",
       color = "Intensity") +
  theme_minimal() +
  theme(legend.position = "right")


```

As can be seen from the above graph, there does seem to be a positive correlation between past cumulative returns and future performance. However, not so straight forward, as there are some funds that had a poor past cumulative return but had good future returns. Lets do some further analysis. 

### correlation analysis between cumulative past returns (3 years) and total flow

```{r}
# Calculate correlation coefficient
correlation_result <- cor.test(combined_analysis$cumulative_return, combined_analysis$total_flow, method = "pearson")

# Print the results
correlation_result

#Creating a nice table for the above results 
correlation_results_df <- data.frame(
  Statistic = c("Correlation Coefficient", "P-Value", "95% Confidence Interval Lower Bound", "95% Confidence Interval Upper Bound"),
  Value = c(0.06977338, 0.04186, 0.002581492, 0.136338075)
)


kable(correlation_results_df, 
      col.names = c("Statistic", "Value"), 
      caption = "Correlation Analysis Results", 
      align = 'c')
```

The above correlation analysis provides some interesting results. The output suggests that there is a weak positive relationship between funds past 3 years performance and fund flows. Thus, although three years past performance may induce some flows towards that fund this pull is not relatively strong. Since the p-value is below 0.05 this weak postie relationship is not random. Therefore, although there is some truth in saying a 3 year strong performing fund will induce more investment. there are most probably a lot more factors that goes into the selection process of investors when deciding where to allocate there funds. 


## Lets create a linear regression to see if 3 years past performance does indicate future performance. 

```{r}
# Linear Regression
model <- lm(future_return ~ cumulative_return, data = combined_analysis)

# Summary of the model
summary(model)

#Making the model tidy and formatting nice
# Tidy the model summary
tidy_model <- tidy(model)

# Create a table with knitr::kable
kable(tidy_model, 
      caption = "Linear Regression Model Summary", 
      align = 'c', 
      digits = 4)

```

This regression analysis shows that there is a statistically significant relationship between past return values and future return values. Thus funds with higher past returns tend to have higher future returns. However the R-squared value is fairly low at 0.12 which suggests that past performance only explain 12% of the variation in future performance. Thus although past performance does have a statistically significant positive relationship with future performance, this relationship is fairly week and should not be solely relied upon.

## Lets evaluate the 10 year look back period 

```{r}
# Define look-back and future performance periods
look_back_years_10 <- 10
future_performance_years <- 1

# Create year column for easier calculations
asisa_data <- asisa_data %>%
  mutate(year = year(date))
```

### Calculating cumulative returns for the 10 year look back period

The cumulative returns of the 10 year period is calculated so that a comparative analysis can be done. 

```{r}
# Calculate cumulative returns for each fund over the look-back period
cumulative_returns_10 <- asisa_data %>%
  group_by(Fund) %>%
  filter(year >= max(year) - look_back_years_10) %>%
  summarize(cumulative_return_10 = prod(1 + Returns) - 1) %>%
  ungroup()

# Rank funds based on these returns
cumulative_returns_10 <- cumulative_returns_10 %>%
  arrange(desc(cumulative_return_10)) %>%
  mutate(rank = row_number())

```


### Analyizing fund flows and future performance

Calculating the future performance (the performance a year after the look back period) will provide somewhat of an answer to whether the look back period was a good indicating of the following period. 

```{r}
# Merge the rankings back to the main data
asisa_data <- asisa_data %>%
  left_join(cumulative_returns_10, by = "Fund")

# Analyze fund flows for the year following the look-back period
fund_flows_analysis_10 <- asisa_data %>%
  filter(year == max(year) - look_back_years_10 + future_performance_years) %>%
  group_by(Fund) %>%
  summarize(total_flow = sum(Flows))

# Calculate future performance
future_performance_10 <- asisa_data %>%
  filter(year >= max(year) - look_back_years_10 + future_performance_years,
         year < max(year) - look_back_years_10 + future_performance_years + future_performance_years) %>%
  group_by(Fund) %>%
  summarize(future_return_10 = prod(1 + Returns) - 1) %>%
  ungroup()

```

### Analyizing data 

To analyze the data, a scatter plot, correlation analysis and regression analysis will be conducted. 

```{r}
# Combine flow analysis and future performance
combined_analysis <- fund_flows_analysis_10 %>%
  left_join(future_performance_10, by = "Fund") %>%
  left_join(cumulative_returns_10, by = "Fund")
```

### Scatter plot illustrating the relationship between past performance and future performance

```{r Figure 2, warning =  FALSE, fig.align = 'center', fig.cap = "10 year look back scatter plot \\label{Figure2}", fig.height = 6, fig.width = 6, dev = 'png'}

# Create a new variable for coloring
combined_analysis <- combined_analysis %>%
  mutate(intensity = cumulative_return_10 + future_return_10)

# Plot with color 
ggplot(combined_analysis, aes(x = cumulative_return_10, y = future_return_10, color = intensity)) +
  geom_point() +
  scale_color_gradient(low = "blue", high = "red") +  # Adjust colors as needed
  labs(title = "Past vs Future Performance of Funds",
       x = "Past Cumulative Return",
       y = "Future Return",
       color = "Intensity") +
  theme_minimal() +
  theme(legend.position = "right")


```

As can be seen from the above figure, there is a far greater variability when the look back period is increased to 10 years. More importantly there still seems to be a positive relationship between past cumulative returns and future returns. However, to identify whether this relationship is stronger or weaker at longer look back periods we will have to complete a correlation and regression analysis. 

### correlation analysis between cumulative past returns (10 years) and total flow

```{r}
# Calculate correlation coefficient
correlation_result_10 <- cor.test(combined_analysis$cumulative_return_10, combined_analysis$total_flow, method = "pearson")

# Print the results
correlation_result_10

# Creating a data frame with the new correlation results
correlation_results_10_df <- data.frame(
  Statistic = c("Correlation Coefficient", "P-Value", "95% Confidence Interval Lower Bound", "95% Confidence Interval Upper Bound"),
  Value = c(0.08477276, 0.08961, -0.01314346, 0.18107861)
)

# Using kable to create a formatted table
kable(correlation_results_10_df, 
      col.names = c("Statistic", "Value"), 
      caption = "Correlation Analysis Results", 
      align = 'c')

```

Like that of the 3 year look back period the correlation between past performance and funds flow is still weakly positive at 0.08. This is stronger than that of the three year look back period at 0.07 however it is not aas strong as I would have thought. Furthermore, the p-value has increased from 0.04 (for the 3 year look back period) to 0.08 which suggests that the results are not statistically significant at the 5% level. However, these results are still statistically significant at the 10% level. The main takeaway from this correlation analysis is that by increasing the look back period we have not increased the correlation between past fund performance and funds flow. 

Lets now have a look at the regression analysis for the 10 year look back period. 


### Lets create a linear regression to see if103 years past performance does indicate future performance. 

```{r}
# Linear Regression
model_10 <- lm(future_return_10 ~ cumulative_return_10, data = combined_analysis)

# Summary of the model
summary(model_10)

#Making the model tidy and formating nice
# Tidy the model summary
tidy_model_10 <- tidy(model_10)

# Create a table with knitr::kable
kable(tidy_model_10, 
      caption = "Linear Regression Model Summary", 
      align = 'c', 
      digits = 4)

```

From the above regression analysis the results provide some interesting analysis. Like that of the correlation analysis, the regression analysis states that there is a weak insignificant relationship between past cumulative returns and future returns. Thus, even if a fund performs well for 10 years, this is not a solid indicator that the fund will perform well in the following period. Furthermore, the regression analysis provides an extremely low r-squared value that suggests that past performance does not hold a significant explanation in the future returns of funds. 

### Summary 

For both the three year and ten year look back periods, past performance is not a good indicator at both fund flows and future performance. this suggests that when investors are looking to invest their money they look at other factors rather than just past performance, which according to our analysis is the right approach, as past performance is not a good indicator at future performance. 





\newpage

# References {-}

<div id="refs"></div>


# Appendix {-}

## Appendix A {-}

Some appendix information here

## Appendix B {-}

