---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Question 2: Currency Hedging Analysis"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Austin Byrne"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Stellenbosch University, Stellenbosch" # First Author's Affiliation
Email1: "22582053\\@sun.ac.za" # First Author's Email address

# Author2: "John Smith"
# Ref2: "Some other Institution, Cape Town, South Africa"
# Email2: "John\\@gmail.com"
#CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.

#Author3: "John Doe"
#Email3: "Joe\\@gmail.com"

#CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
#keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
#JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
abstract: |
  In this question I attempt to replicate the results found in a study that focuses on currency hedging. Secondly I complete my own analysis where I find that a hedged portfolio holds more volatility thus risk compared to a unhedged portfolio. 
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
# Note: Include = FALSE implies the code is executed, but not printed in your pdf.
# warning and message = FALSE implies ugly messages and warnings are removed from your pdf.
# These should be picked up when you execute the command chunks (code sections below) in your rmd, not printed in your paper!

# Lets load in example data, and see how this can be stored and later called from your 'data' folder.
if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)

```


<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Introduction \label{Introduction}

The purpose of the question is two fold. The first section of this question involves attempting to replicate the results shown in a study referring to, "currency hedging - and that there is a
paradox in volatility in that negatively correlated assets may produce portfolio volatilities that
are lower than the sum of its parts". I am unable to fully replicate the findings of the study but i do however plot the scatter plots illustrating the relationship between the ZAR/US exchange rate and a hedged and unhedged portfolio following a portfolio structure of a 60/40 split between equities and bonds and a 70/30 split between global and local.  

The second part of this question I do my own analysis involving the portfolio above and study the volatility comparison of the hedged and unhedged portfolio> In this section I illustrate how the hedged portfolio has far greater volatility than then unhedged portfolio. Thus, which leads the case for not applying long-term (systematic) currency hedging. 

## Loading relevant data

The relevant data is data involves the monthly returns for, MSCI_ACWI, Bbg_Agg, J433, ALBI and the ZAR/USD exchange rate. 

```{r}
indexes <- readRDS("C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/data/Cncy_Hedge_Assets.rds")

ZAR <- readRDS("C:/Users/austi/OneDrive/Desktop/Masters/Financial Econometrics/22582053 (Fin_metrics)/data/Monthly_zar.rds")
```

## Data preperation 

To prepare the data I first merge the monthly returns with the exchange rate data frame> I then find that there are some NA's present which I deal with accordingly. With respect to the NA's present in the values column I fill with the last available value, Which I feel is the most appropriate manner to deal with the NA's. I then also convert the capped SWIX(J433) and ALBI from ZAR to USD for ease of calculation. 

```{r}
# Merge the datasets on date
hedging_data_merged <- left_join(indexes, ZAR, by = "date")

#There are a few NA's in the value column of out merged data set which we need to fix
# Fill NAs with the last available value
hedging_data_merged <- hedging_data_merged %>%
  fill(MSCI_ACWI, Bbg_Agg, J433, ALBI, value, .direction = "down")
#This seems to have fixed our problem so we move on. 

#There are also NA's in the tickers column which we need to fill
# Replace NA in Tickers column with $ZAR.USD
hedging_data_merged <- hedging_data_merged %>%
  mutate(Tickers = replace_na(Tickers, "$ZAR.USD"))

#Lets now convert the capped SWIX(J433) and ALBI from ZAR to USD for ease of calculation
# Convert ZAR returns to USD
hedging_data_merged$J433_USD <- hedging_data_merged$J433 * hedging_data_merged$value
hedging_data_merged$ALBI_USD <- hedging_data_merged$ALBI * hedging_data_merged$value


```

## Lets now try create the portfolio 

### Portfolio weights

I create the weights for the portfolio using the portfolio construction stated in the question. A 60/40 split between equity and bonds, and a 70/30 split between local and global. 

```{r}
#Lets create the weights 
weights <- c(MSCI_ACWI = 0.42, # 60% equity * 70% global
             Bbg_Agg = 0.12,   # 40% bond * 30% global
             J433 = 0.18,      # 60% equity * 30% local
             ALBI = 0.28)      # 40% bond * 70% local
```

### Unhedged and hedged portfolio construction

Here I calculate the returns of the hedged and unhedged portfolios by multiply the specific returns with the accompanying weight vector

```{r}
hedging_data_merged <- hedging_data_merged %>%
  rowwise() %>%
  mutate(
    portfolio_return_unhedged = sum(c_across(c(MSCI_ACWI, Bbg_Agg, J433, ALBI)) * weights),
    portfolio_return_hedged = sum(c_across(c(MSCI_ACWI, Bbg_Agg, J433_USD, ALBI_USD)) * weights)
  ) %>%
  ungroup()

```


## Attempted Study replication 

Within this section of the question, I attempt to replicate the results of the study, although I am not able to fully replicate the paper, I am able to create some interesting scatter plots where I evaluate the relationship between the hedged and unhedged portfolios with the ZAR/USD exchange rate. 


### Unhedged scatter plot 

The unhedged scatter plot illustrates a sightly negative correlation between the ZAR/USD exchange rate and the unhedged portfolio return. This is inline with the findings of the study mentioned in the question. 

```{r Figure 1, warning =  FALSE, fig.align = 'center', fig.cap = "Unhedged vs ZAR/USD scatter plot \\label{Figure1}", fig.height = 6, fig.width = 6, dev = 'png'}

# Scatter plot for Unhedged Portfolio vs ZAR
replication_unhedged_plot <- ggplot(hedging_data_merged, aes(x = value, y = portfolio_return_unhedged)) +
  geom_point() +
  labs(title = "Scatter Plot: ZAR Value vs Unhedged Portfolio Return",
       x = "ZAR Value",
       y = "Unhedged Portfolio Return") +
  theme_minimal()

replication_unhedged_plot
```

### Hedged scatterplot

The findings of the hedged portfolio like that of the unhedged portfolio, find a negative relationship between the ZAR/USD exchange rate and the hedged portfolio returns. Which again is inline with the findings of the study mentioned in the question. 

```{r Figure 2, warning =  FALSE, fig.align = 'center', fig.cap = "hedged vs ZAR/USD scatter plot \\label{Figure2}", fig.height = 6, fig.width = 6, dev = 'png'}

replication_hedged_plot <- ggplot(hedging_data_merged, aes(x = value, y = portfolio_return_hedged)) +
  geom_point() +
  labs(title = "Scatter Plot: ZAR Value vs hedged Portfolio Return",
       x = "ZAR Value",
       y = "hedged Portfolio Return") +
  theme_minimal()

replication_hedged_plot
```

From the analysis completed above it is evident that there is a case for not applying long-term (systematic) currency hedging to your portfolio. To further unpack this statement I dive into my own analysis. 


## Own study

In this section of the question I dive into my own study. I evaluate a comparison between the volatility of a hedge vs unhedged portfolio. 

## Comparison of Hedged vs Unhedged Portfolio Returns 

TO start of this comparison I first run a line graph showing the returns of the hedged portfolio vs the returns of the unhedged portfolio. The results further strengthen the statements made prior that the hedged portfolio returns induce greater volatility than that of the unhedged portfolio. To further unpack this discussion on volatility comparison I use the PerformanceAnalytics package in R to fully evaluate the volatility of the portfolios in question. 

```{r Figure 3, warning =  FALSE, fig.align = 'center', fig.cap = "hedging comparisson plot \\label{Figure3}", fig.height = 6, fig.width = 6, dev = 'png'}

pacman::p_load(fmxdat)

hedging_comparison_plot <- hedging_data_merged %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = portfolio_return_unhedged, color = "Unhedged Portfolio"), size = 1) +
  geom_line(aes(y = portfolio_return_hedged, color = "Hedged Portfolio"), size = 1)  +
    
  fmxdat::theme_fmx(title.size = ggpts(20), 
            subtitle.size = ggpts(15),
            caption.size = ggpts(12) )+
  scale_color_manual(values = c("Unhedged Portfolio" = "blue", "Hedged Portfolio" = "orange")) +
    
    fmxdat::fmx_cols() + 
    
    labs(x = "", y = "%", caption = "Note:\nCalculation own",
       title = "Comparison of Hedged vs Unhedged Portfolio Returns",
       subtitle = "Tracking the Performance Over Time")

#for finishing touches use finplot
fmxdat::finplot(hedging_comparison_plot,x.vert = T, x.date.type = "%Y", x.date.dist = "1 years", )

```

## Understanding the volatility  of the hedged vs unhedged portfolios

In this section I calculate the rolling volatility and plot the results for the hedged vs unhedged portfolios. To further the analysis I evaluate a rolling period of 1 year with one of 3 years. 

### Calculating the rolling volatility

Using the performance analytics package in R I am able to calculate the 1 year rolling volatility. 

```{r}
#loading the performance analytics package
pacman::p_load(PerformanceAnalytics)
pacman::p_load(TTR)
pacman::p_load(tbl2xts)

#The performance analytics package requires the data to be in xts format
#I tried using the tbl2xts package here to convert to xts but for some reason I could not get it to work, I therefore had to find another way.
hedging_xts_unhedged <- xts(hedging_data_merged$portfolio_return_unhedged, order.by = as.Date(hedging_data_merged$date))
hedging_xts_hedged <- xts(hedging_data_merged$portfolio_return_hedged, order.by = as.Date(hedging_data_merged$date))


# Calculate rolling volatility (standard deviation)
roll_vol_unhedged <- runSD(hedging_xts_unhedged, n = 12) # Annualized


roll_vol_hedged <- runSD(hedging_xts_hedged, n = 12) # Annualized

```

### Rolling Realized Volatility: Hedged vs Unhedged Portfolio plot

Here I plot a comparison of the rolling 1 year realized volatility of the hedged vs unhedged portfolio using a line graph. As expected the hedged portfolio continues to be accompanied by higher volatility compared to that of the unhedged portfolio. To evaluate whether this is just the case with the 1 year rolling volatility I evaluate the 3 year rolling volatility next. 

```{r}
# Creating a combined dataframe for easy plotting
# Since I could not get the tbl2xts package to work I had to find a new method to get my data out of the xts format to a nice data frame where I can now do some plots. 
hedging_vol_data <- data.frame(
  date = index(roll_vol_unhedged),
  roll_vol_unhedged = as.numeric(roll_vol_unhedged),
  roll_vol_hedged = as.numeric(roll_vol_hedged)
)

#Plotting

 hedging_vol_data_plot <- hedging_vol_data %>%
  ggplot() + 
  geom_line(aes(date, roll_vol_unhedged, color = "Unhedged"), size = 2, alpha = 0.7) +
  geom_line(aes(date, roll_vol_hedged, color = "Hedged"), size = 2, alpha = 0.7) +
  
  
  fmxdat::theme_fmx(title.size = fmxdat::ggpts(30), 
                    subtitle.size = fmxdat::ggpts(28),
                    caption.size = fmxdat::ggpts(25),
                    CustomCaption = TRUE) + 
  
 
  fmxdat::fmx_cols() + 
  
  labs(x = "", y = "Volatility", 
       color = "Portfolio Type",
       caption = "Note: Own calculation",
       title = "Rolling Realized Volatility: Hedged vs Unhedged Portfolio",
       subtitle = "Comparing the impact of hedging on portfolio volatility") 
  
  # Finishing touches
  fmxdat::finplot(hedging_vol_data_plot, x.vert = TRUE, x.date.type = "%Y", x.date.dist = "1 years")


```


## Lets now consider a longer rolling term of three years and evaluate its impact on valatility

Now considering a longer rolling term of three years I evaluate whether the hedged portfolio still contains higher volatility compared to that of the unhedged portfolio. 


### Calculating rolling volatility using 36 months

I calculate the 3 year rolling volatility for both the hedged and unhedged portfolios. 

```{r}
# Calculate rolling volatility (standard deviation)
roll_vol_unhedged_36 <- runSD(hedging_xts_unhedged, n = 36)


roll_vol_hedged_36 <- runSD(hedging_xts_hedged, n = 36) 
```

### 3 year rolling realized volatility: Hedged vs Unhedged Portfolio plot

By utilizing the same methods as was used in calculating the 1 year rolling realized volatility I am able to calculate the three year rolling realized volatility and plot the results. 

Once again it is evident that the hedged portfolio contains higher volatility and therefore risk than that of the unhedged portfolio. The 3 year rolling return results in lower volatility than that of the 1 year rolling volatility for the portfolios which can be expected due to increased time period.

```{r}
# Creating a combined dataframe for easy plotting
# Since I could not get the tbl2xts package to work I had to find a new method to get my data out of the xts format to a nice data frame where I can now do some plots. 
hedging_vol_data_36 <- data.frame(
  date = index(roll_vol_unhedged_36),
  roll_vol_unhedged_36 = as.numeric(roll_vol_unhedged_36),
  roll_vol_hedged_36 = as.numeric(roll_vol_hedged_36)
)

#Plotting

 hedging_vol_data_plot_36 <- hedging_vol_data_36 %>%
  ggplot() + 
  geom_line(aes(date, roll_vol_unhedged_36, color = "Unhedged"), size = 2, alpha = 0.7) +
  geom_line(aes(date, roll_vol_hedged_36, color = "Hedged"), size = 2, alpha = 0.7) +
  
  
  fmxdat::theme_fmx(title.size = fmxdat::ggpts(30), 
                    subtitle.size = fmxdat::ggpts(28),
                    caption.size = fmxdat::ggpts(25),
                    CustomCaption = TRUE) + 
  
 
  fmxdat::fmx_cols() + 
  
  labs(x = "", y = "Volatility", 
       color = "Portfolio Type",
       caption = "Note: Own calculation",
       title = "3 year rolling Realized Volatility: Hedged vs Unhedged Portfolio",
       subtitle = "Comparing the impact of hedging on portfolio volatility") 
  
  # Finishing touches
  fmxdat::finplot(hedging_vol_data_plot_36, x.vert = TRUE, x.date.type = "%Y", x.date.dist = "1 years")

```

## Conclusion 

Thus, from the analysis completed above it  is evident that the hedging your portfolio to reduce risk may not be the appropriate way at tackling risk. Hedging your portfolio may actually induce more volatility resulting in higher risk. 




\newpage

# References {-}

<div id="refs"></div>


# Appendix {-}

## Appendix A {-}

Some appendix information here

## Appendix B {-}

